Cloud Computing~\cite{mell2011nist} systems and Service Oriented Architectures (SOA) have 
reached a level of complexity~\cite{Huebscher:2008:SAC:1380584.1380585,Conejero:2012:MSQ:2357487.2357591} 
that implies the necessity of new methods and algorithms to automatically deal 
with the vast amount of data, variables, parameters, etc. that appears 
in this new realm for the advanced management of applications, services or resources. 

In this new environment QoS is playing a relatively minor role but its 
importance, in a wide range of applications scenarios, is likely become more 
crucial than ever before. The main problem lies in the complexity 
of designing QoS models that enables an adequate management of a distributed 
architecture making decisions about resource provisioning, getting feedback for 
the final users, etc. with the objective of avoiding existing ``brute-force''
solutions and overprovisioning. 

Although QoS management and the concept of ``Measured Service''~\cite{mell2011nist} have 
been widely investigated~\cite{Conejero:2012:MSQ:2357487.2357591} in the well-known grid-computing area, 
the emergence of the Cloud Computing paradigm brings a new set of open issues: accomplish the 
Service Level Agreements (SLAs), predict future workload, process large and diverse data streams/logs, 
make real-time decisions, reasoning and inference, dynamic adaptation and provision of 
resources, etc. Nowadays old-fashioned QoS management techniques (based on a combination of network resource 
provisioning with techniques such as admission control or active queue management) can be applied to a static 
environment but in the near future with the new implications for professionals~\cite{DBLP:journals/jucs/Colomo-PalaciosFSS12}, the challenge of 
providing higher elasticity and dynamic adaptation cannot be accomplished with these methods. 

In this sense, the features and requirements of the new cloud systems with regards 
to QoS~\cite{Pedersen:2011:AMQ:2114495.2115542} and the aforementioned implications in the 
future IT~\cite{DBLP:journals/jucs/Colomo-PalaciosFSS12} match the advantages 
of software component and knowledge-based architectures. In fact, Autonomic Computing support 
for the next generation of cloud systems needs to be~\cite{Conejero:2012:MSQ:2357487.2357591,Pedersen:2011:AMQ:2114495.2115542}: 
1) Self-x management, 2) agile, flexible and reliable, 4) deployable over a multiple cloud platforms, 5) handle complexity, 6) enable 
collaboration and coordination and 7) cost-effective and greener 
(energy-efficient). Under this context, semantic technologies have emerged as an 
option to design and develop intelligent software components and agents to 
perform certain tasks on the Web and fulfill user's requirements (in this case 
applications). Therefore, Semantics enables machines to automatically process 
and enrich data from different sources and has the potential to deeply influence 
the further development of the Internet Economy as cloud systems also does.

On the other hand, in the Semantic Web area there is a growing commitment to process large data streams applying 
new stream reasoning~\cite{Bolles:2008:SSE:1789394.1789438,Barbieri:2010:EEC:1739041.1739095} 
or complex event processing~\cite{Anicic:2011:EUL:1963405.1963495} (CEP) techniques. Furthermore there are research works offering cloud-based 
solutions to deal with Big Data~\cite{Fan:2013:MBD:2481244.2481246} (e.g. analysis of social media), modeling SLAs and ECA rules with ontologies, 
monitoring real-time systems (e.g. traffic), sensor networks, or making decisions in a collaborative fashion~\cite{RodriGuez-GonzaLez:2012:UAP:2350799.2350907} (e.g. clinical reasoning). 
The main advantage of applying semantic technologies to a specific domain lies in the standard representation of knowledge and data through a common-shared data model (RDF) and the 
capacity of reusing existing knowledge through ontologies (OWL). Thus, data coming from cloud systems can be automatically processed, checked for inconsistencies and 
used in expert systems to support self-x management activities.

This review is intended to provide researchers, developers and practitioners a summary of the current status of QoS management in Cloud Computing and SOA applying semantics. To do so, the paper 
is structured as follows. Next section reviews the background concepts required to a better understanding of the paper. Section~\ref{qos-semantics} 
presents the existing works to perform QoS management using semantics; more specifically most of the ontology-based frameworks for 
QoS management are reviewed. Afterwards, a review of existing techniques for processing large data streams is also provided in Section~\ref{data-stream}. 
Section~\ref{framework} outlines a framework to meet QoS requirements applying semantics and data stream processing techniques in real-time. 
Finally, the paper ends with a discussion of existing approaches for semantic-based QoS management, 
limitations, future challenges and concluding remarks. 

